---
title: Association rule mining to find correlations between presidential preference and county-level population demographics using Census data
author: "Danny Morris"
date: "12/3/2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## R packages

```{r}
library(dplyr)
library(tibble)
library(tidycensus)
library(arules)
library(DT)
```

## Load functions and Census API key

```{r}
census_api_key(Sys.getenv("CENSUS_API_KEY"))

source("R/utils/arules_utils.R")
source("R/voting_results.R")
source("R/acs/population.R")
source("R/acs/age.R")
source("R/acs/race.R")
source("R/acs/gini.R")
source("R/acs/income.R")
source("R/acs/unemployment.R")
source("R/acs/education.R")
source("R/acs/industry.R")
source("R/acs/citizenship.R")
source("R/acs/insurance.R")
source("R/acs/veterans.R")
```

## Data

### Presidential voting outcomes

```{r}
votes <- get_voting_results()

datatable(votes) %>%
  formatRound(columns = c("trump_pct", "biden_pct", "margin"), digits = 3)
```

### American Community Survey (ACS) demographic variables

County-level demographic characteristics taken from the 2018 5-year American Community Survey for use in the association rules analysis:

- Population counts by age band
- Population counts by race
- Gini index (income inequality)
- Income per capita
- Unemployment rate
- Population counts by educational attainment
- Population counts by industry
- Population counts by citizenship status
- Population counts by insurance coverage
- Population counts by veteran status

```{r}
population <- get_population(tag = "POPULATION")
age <- get_age(tag = "AGE")
race <- get_race(tag = "RACE")
gini <- get_gini(tag = "GINI")
income <- get_income(tag = "INCOME")
unemployment <- get_unemployment(tag = "UNEMPLOY")
education <- get_education(tag = "EDU_ATTAIN")
industry <- get_industry(tag = "INDUSTRY")
citizenship <- get_citizenship(tag = "CITIZEN")
insurance <- get_insurance(tag = "HEALTH_INSURANCE")
veterans <- get_veterans(tag = "VETERANS")
```

### Merge votes and ACS data

```{r}
data_joins <- votes %>%
  select(fips, state, name, party_winner, trump_pct, margin) %>%
  inner_join(population, by = c("fips" = "GEOID")) %>%
  inner_join(age, by = c("fips" = "GEOID")) %>%
  inner_join(race, by = c("fips" = "GEOID")) %>%
  inner_join(gini, by = c("fips" = "GEOID")) %>%
  inner_join(income, by = c("fips" = "GEOID")) %>%
  inner_join(unemployment, by = c("fips" = "GEOID")) %>%
  inner_join(education, by = c("fips" = "GEOID")) %>%
  inner_join(industry, by = c("fips" = "GEOID")) %>%
  inner_join(citizenship, by = c("fips" = "GEOID")) %>%
  inner_join(insurance, by = c("fips" = "GEOID")) %>%
  inner_join(veterans , by = c("fips" = "GEOID")) %>%
  drop_na()

colnames(data_joins) %>%
  enframe(name = "column_position", value = "column_name") %>%
  datatable(caption = "Column names")
```

### Adjust demographics by population size

```{r}
population_adjusted <- data_joins %>%
  mutate_at(vars(AGE_18_29,
                 AGE_30_44,
                 AGE_45_59,
                 AGE_60_Plus,
                 contains("UNEMPLOY"),
                 contains("HEALTH_INSURANCE"),
                 contains("VETERANS"),
                 contains("EDU_ATTAIN"),
                 contains("INDUSTRY")), 
            list(~./AGE_18_Plus)) %>%
  mutate_at(vars(contains("RACE"),
                 contains("CITIZEN")),
            list(~./POPULATION_Total)) 
```

## Association rule mining

### Convert numeric variables to discrete categories

```{r}
decile_data <- population_adjusted %>%
  select(-fips, -state, -name, -party_winner, -margin, 
         -AGE_18_Plus, -POPULATION_Total) %>%
  mutate(trump_pct = bucket_percentage(trump_pct)) %>%
  mutate_at(vars(-trump_pct), convert_numeric_to_discrete) 
```

### Convert discrete categories to binary indicators

```{r}
binary_data <- decile_data %>%
  convert_to_dummies()
```

### Convert binary indicators to transactions

```{r}
transactions <- convert_df_to_transactions(df = binary_data)
```

### Generate association rules for strong Trump support

```{r}
trump_arules <- run_apriori(
  trans = transactions,
  min_supp = 0.001,
  min_conf = 0.001,
  minlen = 2,
  maxlen = 2,
  default = "lhs",
  rhs = "trump_pct_X4"
)

trump_arules %>%
  convert_rules_to_df() %>%
  make_rules_DT()
```

### Generate association rules for strong Biden support

```{r}
biden_arules <- run_apriori(
  trans = transactions,
  min_supp = 0.001,
  min_conf = 0.001,
  minlen = 2,
  maxlen = 2,
  default = "lhs",
  rhs = "trump_pct_X1"
)

biden_arules %>%
  convert_rules_to_df() %>%
  make_rules_DT()
```


