---
title: "Exploratory Data Mining: Collecting county-level voting and demographic data"
author: "Danny Morris"
date: "12/3/2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r, eval=F, echo=F}
# set the working directory for interactive development
# but ignore when knitting the document
setwd("get-data")
```

## R packages

```{r}
# install.packages("tidycensus")
# install.packages("gistr")
# install.packages("DT")
library(dplyr)
library(tibble)
library(tidycensus)
library(DT)
library(gistr)
```

## Load functions and Census API key

```{r}
census_api_key(Sys.getenv("CENSUS_API_KEY"))

source("R/voting_results.R")
source("R/acs/population.R")
source("R/acs/age.R")
source("R/acs/race.R")
source("R/acs/gini.R")
source("R/acs/income.R")
source("R/acs/unemployment.R")
source("R/acs/education.R")
source("R/acs/industry.R")
source("R/acs/citizenship.R")
source("R/acs/insurance.R")
source("R/acs/veterans.R")
```

## Data

### Presidential voting outcomes

```{r}
votes <- get_voting_results(filename="presidential.csv")

datatable(votes)
```

### American Community Survey (ACS) demographic variables

County-level demographic characteristics taken from the 2018 5-year American Community Survey for use in the association rules analysis:

- Population counts by age band
- Population counts by race
- Gini index (income inequality)
- Income per capita
- Unemployment rate
- Population counts by educational attainment
- Population counts by industry
- Population counts by citizenship status
- Population counts by insurance coverage
- Population counts by veteran status

```{r, cache=T}
population <- get_population(tag = "POPULATION")
age <- get_age(tag = "AGE")
race <- get_race(tag = "RACE")
gini <- get_gini(tag = "GINI")
income <- get_income(tag = "INCOME")
unemployment <- get_unemployment(tag = "UNEMPLOY")
education <- get_education(tag = "EDU_ATTAIN")
industry <- get_industry(tag = "INDUSTRY")
citizenship <- get_citizenship(tag = "CITIZEN")
insurance <- get_insurance(tag = "HEALTH_INSURANCE")
veterans <- get_veterans(tag = "VETERANS")
```

### Merge votes and ACS data

```{r}
data_joins <- votes %>%
  select(fips, state, name, party_winner, trump_pct, margin) %>%
  inner_join(population, by = c("fips" = "GEOID")) %>%
  inner_join(age, by = c("fips" = "GEOID")) %>%
  inner_join(race, by = c("fips" = "GEOID")) %>%
  inner_join(gini, by = c("fips" = "GEOID")) %>%
  inner_join(income, by = c("fips" = "GEOID")) %>%
  inner_join(unemployment, by = c("fips" = "GEOID")) %>%
  inner_join(education, by = c("fips" = "GEOID")) %>%
  inner_join(industry, by = c("fips" = "GEOID")) %>%
  inner_join(citizenship, by = c("fips" = "GEOID")) %>%
  inner_join(insurance, by = c("fips" = "GEOID")) %>%
  inner_join(veterans , by = c("fips" = "GEOID")) %>%
  drop_na()

data_joins %>%
  head() %>%
  datatable() %>%
  formatRound(
    columns = c("trump_pct", "margin"), 
    digits = 3
  )
```

### Adjust demographics by population size

```{r}
population_adjusted <- data_joins %>%
  mutate_at(vars(AGE_18_29,
                 AGE_30_44,
                 AGE_45_59,
                 AGE_60_Plus,
                 contains("UNEMPLOY"),
                 contains("HEALTH_INSURANCE"),
                 contains("VETERANS"),
                 contains("EDU_ATTAIN"),
                 contains("INDUSTRY")), 
            list(~./AGE_18_Plus)) %>%
  mutate_at(vars(contains("RACE"),
                 contains("CITIZEN")),
            list(~./POPULATION_Total)) 

population_adjusted %>%
  head() %>%
  datatable() %>%
  formatRound(
    columns = 5:52, 
    digits = 3
  )
```

## Write data to local file

```{r, eval=F}
write_csv(population_adjusted, "county_data.csv")
```

## Session info

```{r}
sessionInfo()
```

